---
title: "INFO 550 Final Project- Shaily Krishan"
output:
  html_document:
    df_print: paged
---

```{r}
## Install the required package with:
## install.packages("RSocrata")
## Reference: https://dev.socrata.com/foundry/data.medicare.gov/eqxu-aw4f

library("RSocrata")

# Reading in the "Dialysis Facility Compare"" data
dfc.df <- read.socrata(
  "https://data.medicare.gov/resource/eqxu-aw4f.json",
  app_token = "a2pVgaK9QKjxN5iNc4DLohtjI",
  email     = "shells.27@gmail.com",
  password  = "MCh@nge2018",
  stringsAsFactors = FALSE
)
```
# Reading in "ESRD QIP - Complete QIP Data - Payment Year 2018" data
```{r}
esrdqip.df <- read.socrata(
  "https://data.medicare.gov/resource/m26r-4ib8.json",
  app_token = "a2pVgaK9QKjxN5iNc4DLohtjI",
  email     = "shells.27@gmail.com",
  password  = "MCh@nge2018",
  stringsAsFactors = FALSE
)
```
# Reading in "Patient survey (ICH CAHPS)" data
```{r}
patsurvey.df <- read.socrata(
  "https://data.medicare.gov/resource/rza3-mnpj.json",
  app_token = "a2pVgaK9QKjxN5iNc4DLohtjI",
  email     = "shells.27@gmail.com",
  password  = "MCh@nge2018",
  stringsAsFactors = FALSE
)
```
# Create dataframe with selected DFC variables
```{r}
library(dplyr)
dfc <- dfc.df %>% select(provider_number,facility_name,network,address_line_1,
                         city,state,zip,chain_owned,five_star,
                         X_of_dialysis_stations,mortality_rate_facility,
                         readmission_rate_facility,
                         standardized_hospitalization_ratio,
                         standardized_readmission_ratio,
                         standard_infection_ratio)
```
# Create dataframe with selected ESRDQIP variables
```{r}
library(dplyr)
esrdqip <- esrdqip.df %>% select(cms_certification_number_ccn,
                                 strr_measure_score,total_performance_score,
                                 nhsn_bsi_measure_score,
                                 kt_v_dialysis_adequacy_combined_measure_score,
                                 anemia_management_measure_score,
                                 vascular_access_combined_measure_score,
                                 mineral_metabolism_measure_score)

```
# Create dataframe with selected Patient survey variables
```{r}
patsurvey <- patsurvey.df %>% select(provider_number, starts_with("top_box"))
names(patsurvey) <- gsub("top_box_percent_of_patients_","",names(patsurvey))
```
# Reading in Dialysis Facility Report file
```{r}
library(readr)
setwd("~/INFO 550/Project/Files")
dfr <- read_delim("DFR_Data_FY2017.csv", delim = ",")

# Selecting demographic, ESRD cause, Transplant variables from DFR
fac_demo <- select(dfr,provfs,rac1y4_f,
                   rac2y4_f,rac3y4_f,rac4y4_f,rac5y4_f,eth1y4_f,
                   eth2y4_f,eth3y4_f,sexy4_f,agey4_f,dis1y4_f,dis2y4_f,
                   dis3y4_f,dis4y4_f,rdsallty4_f,txally4_f)

# Create data frame from DFR variables
fac_demo1 <- data.frame(fac_demo)
```
# Join data frames
```{r}
# DFC + DFR data
df1 <- left_join (dfc,fac_demo1, by=c("provider_number"="provfs"))
# ESRDQIP + Patsurvey data
df2 <- left_join(patsurvey,esrdqip,
                 by=c("provider_number"="cms_certification_number_ccn"))
# DFC + DFR + ESRDQIP + Patsurvey data
allfacdata <- left_join(df1,df2,by=c("provider_number"="provider_number"))

# Conversion of rate & measure variables to numeric
# Columns 9 through 44 are rate/ measure variables which will be converted to numeric, columns 1 though 8 are retained as character variables)
allfac <- data.frame(allfacdata[,1:8], sapply(allfacdata[,9:44], function(x) as.numeric(x)))
```
# Adding zipcode xy coordinates 
```{r}
library(dplyr)
# Install "zipcode" package
library(zipcode)
data("zipcode")
# Select columns from zipcode 
zipcoord <- zipcode %>% select(zip,longitude,latitude)
# Join zip code data
geo_join <- left_join(allfac, zipcoord, by=c("zip"="zip"))
```
# Using SQLite to import CKD prevalence data
```{r}
# Set working directory to location of database
setwd("C:/Users/shail")
# load libraries
library(DBI)
library(RSQLite)

con <- dbConnect(RSQLite::SQLite(),dbname="CKD_Prevalence.db")
# Show tables in database
dbListTables(con)
# Show types of prevalence measures
dbGetQuery(con,"select distinct DataValueTypeID from CKD_prev")
# Get Crude prevalence values for each state into a dataframe
ckd_prev <- dbGetQuery(con,"select LocationAbbr,avg(DataValue) from CKD_prev 
                       where DataValueTypeID='CrdPrev' and 
                       Question like '%kidney%' and 
                       YearStart='2016' group by LocationAbbr")
dbDisconnect(con)
```
# Join Prevalence data
```{r}
library(dplyr)
alldata <- left_join(geo_join,ckd_prev,by=c("state"="LocationAbbr"))
```
# Creating categories in continuous variables
```{r}
# Variable 'perfcats' for Facility performance score categories
alldata$perfcats <- cut(alldata$total_performance_score,
                         breaks = quantile(alldata$total_performance_score,na.rm = TRUE),
                         labels=c("Poor","Fair","Good","Very good"),include.lowest=TRUE)

# Variable 'facrating' for patient's rating of Dialysis facilities
alldata$facrating <- cut(alldata$rating_of_the_dialysis_facility,
                         breaks = quantile(alldata$rating_of_the_dialysis_facility,na.rm = TRUE),
                         labels=c("Poor","Fair","Good","Very good"),
                         include.lowest=TRUE)

# Variable 'stations' for number of dialysis stations categories
alldata$stations <- cut(alldata$X_of_dialysis_stations,breaks=c(0,mean(alldata$X_of_dialysis_stations),max(alldata$X_of_dialysis_stations)),c("up to 17","17 to 80"))

# Make 'five_star' a factor 
alldata$five_star <- factor(alldata$five_star)
```
# Exploratory analysis
```{r}
library(ggplot2)
library(dplyr)
# Number of facilities in networks
# Networks 6,14,9 are largest networks- they have top 3 highest number of facilities
alldata %>% group_by(network) %>% summarize(total=n()) %>% arrange(desc(total))
# Number of facilities in networks, by chain status
# Most facilities are chain owned
alldata %>% ggplot(aes(x=network)) + 
  geom_bar(aes(fill=chain_owned),stat="count") + 
  ggtitle("Network facilities by chain status")
# Total performance score in networks
alldata %>% ggplot(aes(x=network)) + 
  geom_bar(aes(fill=perfcats),stat="count") + 
  ggtitle("Total performance scores in network facilities")
summary(alldata$perfcats)
```
# Mortality rate in facilities
```{r}
library(ggplot2)
library(dplyr)
summary(alldata$mortality_rate_facility)
# Histogram of Mortality rates 
alldata %>% ggplot(aes(x=mortality_rate_facility)) + geom_histogram() + 
  ggtitle("Mortality Rates in facilities")

# Scatterplot of Mortality rate vs readmission rate
alldata %>% ggplot(aes(x=mortality_rate_facility,y=readmission_rate_facility)) + geom_point()+geom_smooth(method="lm") + ggtitle("Mortality rate vs readmission rate")

# Boxplot of mortality by performance score categories
ggplot(data = alldata, 
       aes(x = alldata$perfcats, y = alldata$mortality_rate_facility)) + 
    geom_boxplot() + xlab("Facility performance score categories") + 
  ylab("Facility mortality rate")
# Boxplot of mortality rate by five-star category
ggplot(data = alldata, 
       aes(x = alldata$five_star, y = alldata$mortality_rate_facility)) + 
    geom_boxplot() + xlab("Facility five star rating") + 
  ylab("Facility mortality rate")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=readmission_rate_facility,color=five_star)) + geom_point() + ggtitle("Mortality rate vs readmission rate")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=readmission_rate_facility,color=chain_owned)) + geom_point() + ggtitle("Mortality rate vs readmission rate") + facet_wrap(~five_star)

# Filtering higher mortality cases
alldata %>% filter(mortality_rate_facility >= 30) %>% select(chain_owned,network,state)
alldata %>% filter(mortality_rate_facility >= 30) %>% group_by(chain_owned) %>% summarize(n())
```
# CKD prevalence
```{r}
# Summary of prevalence of Chronic Kidney Disease within networks
# Network 10 has highest prevalence
alldata %>% group_by(network) %>% summarize(meanprev=mean(`avg(DataValue)`)) %>% arrange(desc(meanprev))
# Plotting prevalence of CKD in networks (no values for network 17-
# Northern California, Hawaii, Guam, Mariana Islands, and American Samoa- no data
alldata %>% group_by(network) %>% summarize(meanprev=mean(`avg(DataValue)`)) %>% ggplot(aes(x=network,y= meanprev)) + geom_bar(stat = "identity") + 
  ggtitle("Mean Chronic Kidney Disease (CKD) prevalence in networks")
```
# Facility performance scores
```{r}
library(ggplot2)
library(dplyr)
# Histogram of total performance score
alldata %>% ggplot(aes(x=total_performance_score)) + geom_histogram() + ggtitle("Total performance score by star rating") + facet_wrap(~five_star)

# Scatter plots
alldata %>% ggplot(aes(x=standardized_hospitalization_ratio,y=total_performance_score)) + geom_point()+geom_smooth(method="lm") + ggtitle("Hospitalization rate vs Total perf score")

alldata %>% ggplot(aes(x=readmission_rate_facility,y=total_performance_score)) + geom_point()+geom_smooth(method="lm") + ggtitle("Readmission rate vs Total perf score")

alldata %>% ggplot(aes(x=five_star,y=total_performance_score)) + 
    geom_boxplot() + xlab("Facility Five Star Rating") + 
  ylab("Total Performance Score")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=total_performance_score,color=five_star)) + geom_point() + ggtitle("Mortality rate vs total perf score")

alldata %>% group_by(network) %>% summarize(meanscore=mean(total_performance_score,na.rm = TRUE)) %>% arrange(desc(meanscore))

alldata %>% group_by(network) %>% summarize(meanscore=mean(total_performance_score,na.rm = TRUE)) %>% ggplot(aes(x=network,y= meanscore)) + geom_bar(stat = "identity") + 
  ggtitle("Mean facility performance score in networks")
```
# Examining 'poor' performing networks & states
```{r}
# Top 10 states with the highest number of 'poor' performing facilities
alldata %>% filter(perfcats == "Poor") %>% group_by(state) %>% 
    summarize(total=n()) %>% arrange(desc(total)) %>% head(.,10) %>% ggplot(aes(x=state,y=total)) +
  geom_bar(stat="identity") + ggtitle("Top 10 states with highest number of poor preformance facilities")

# Top 10 networks with the highest number of 'poor' performing facilities

alldata %>% filter(perfcats == "Poor") %>% group_by(network) %>% 
    summarize(total=n()) %>% arrange(desc(total)) %>% head(.,10) %>% ggplot(aes(x=network,y=total)) +
  geom_bar(stat="identity") + ggtitle("Top 10 networks with highest number of poor preformance facilities")

# Fewer facilities with higher mortality & readmission rates in good-very good performance
alldata %>% ggplot(aes(x=mortality_rate_facility)) + geom_histogram() + facet_wrap(~perfcats) + ggtitle("Mortality Rates by facility performance scores")


alldata %>% ggplot(aes(x=readmission_rate_facility)) + geom_histogram() + facet_wrap(~perfcats) + ggtitle("Readmission Rates by facility performance scores")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=standardized_hospitalization_ratio)) + geom_point()+geom_smooth(method="lm") + ggtitle("Mortality rate vs SHR")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=standard_infection_ratio)) + geom_point()+geom_smooth(method="lm") + ggtitle("Mortality rate vs SIR")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=standardized_readmission_ratio)) + geom_point()+geom_smooth(method="lm") + ggtitle("Mortality rate vs SRR")

alldata %>% ggplot(aes(x=mortality_rate_facility,y=standardized_readmission_ratio)) + geom_point()+geom_smooth(method="lm") + facet_wrap(~perfcats) + ggtitle("Mortality rate vs SRR by performance scores")
```
# Five star ratings
```{r}
# Majority of facilities are 3,4,5 star rated, 698 facilities are 1 or 2 star rated: FL, TX, GA, CA, NY, OH, IL, TN, LA, MI 
summary(alldata$five_star)
alldata %>% ggplot(aes(x=five_star)) + geom_bar(aes(fill=perfcats),stat="count")
star <- c("1","2")
alldata %>% filter(five_star %in% star) %>% group_by(state) %>% summarize(total=n()) %>% arrange(desc(total)) %>% head(.,10)

alldata %>% filter(perfcats == "Poor") %>% group_by(state) %>% 
    summarize(total=n()) %>% ggplot(aes(x=state,y=total)) + geom_bar(stat = "identity") + ggtitle("States facilities with poor total performance scores")

# Histogram shows that even though fewer facilities have 1 & 2 star ratings, the mortality in these facilities is much higher (40)
alldata %>% ggplot(aes(x=mortality_rate_facility)) + geom_histogram() + facet_wrap(~five_star) + ggtitle("Mortality Rates by Star Rating")
```
# Patient characteristics
```{r}
alldata %>% ggplot(aes(x=dis2y4_f,y=rac2y4_f)) + geom_point()+geom_smooth(method="lm") + ggtitle("AA race vs Hypertension") + facet_wrap(~five_star)
alldata %>% group_by(network) %>% summarize(AA=mean(rac2y4_f,na.rm = TRUE)) %>% ggplot(aes(x=network,y= AA)) + geom_bar(stat = "identity") + 
  ggtitle("Percentage of AA patients in networks")
alldata %>% ggplot(aes(x=dis1y4_f)) + geom_histogram() + ggtitle("Percentage of patients with Diabetes as primary ESRD cause")
alldata %>% ggplot(aes(x=dis2y4_f)) + geom_histogram() + ggtitle("Percentage of patients with Hypertension as primary ESRD cause")
alldata %>% ggplot(aes(x=dis3y4_f)) + geom_histogram() + ggtitle("Percentage of patients with Glomerulonephritis as primary ESRD cause")
alldata %>% ggplot(aes(x=rac2y4_f)) + geom_histogram() + ggtitle("Percentage of AA patients") + facet_wrap(~perfcats)
alldata %>% group_by(network) %>% summarize(AA=mean(rac4y4_f,na.rm = TRUE)) %>% ggplot(aes(x=network,y= AA)) + geom_bar(stat = "identity") + 
  ggtitle("Percentage of white patients in networks")
```
# Transplants
```{r}
summary(alldata$rdsallty4_f)
alldata %>% group_by(network) %>% summarize(meantx=mean(rdsallty4_f,na.rm = TRUE)) %>% ggplot(aes(x=network,y= meantx)) + geom_bar(stat = "identity") + 
  ggtitle("Mean kidney transplants in networks")
alldata %>% ggplot(aes(x=rdsallty4_f)) + geom_histogram() + facet_wrap(~five_star) + ggtitle("Transplants by Star Rating")
alldata %>% ggplot(aes(x=rdsallty4_f)) + geom_histogram() + facet_wrap(~perfcats) + ggtitle("Transplants by performance score")
alldata %>% ggplot(aes(x=rdsallty4_f)) + geom_histogram() + facet_wrap(~stations) + ggtitle("Transplants by number of dialysis stations")
alldata %>% ggplot(aes(x=rdsallty4_f,y=X_of_dialysis_stations)) + geom_point()+geom_smooth(method="lm") + ggtitle("Transplants vs no. of dialysis stations")
```
# Mapping
```{r}
library(ggmap)
library(ggplot2)
usbasemap <- get_map(location = 'United States', zoom= 4, maptype = "terrain", source='google', color='color')
ggmap(usbasemap)
```
# Mapping poor performing facilities
```{r}
library(ggmap)
library(ggplot2)
# Filter cases with "poor" score
poorperf <- alldata %>% filter(perfcats == "Poor")
geopoorperf <- ggmap(usbasemap) + geom_point(aes(x=longitude,y=latitude),data=poorperf, na.rm = T, color="red", alpha=0.3,size=1) + ggtitle("Facilities with lowest performance scores")
# Filter cases with 1 & 2 star rating
star <- c(1,2)
lowstar <- alldata %>% filter(five_star %in% star)
geolowstar <- ggmap(usbasemap) + geom_point(aes(x=longitude,y=latitude),data=lowstar, na.rm = T, color="blue", alpha=0.3,size=1) + ggtitle("Facilities with 1 & 2 star ratings")
plot(geopoorperf)
plot(geolowstar)
```
# Geographical distribution of poor performing states
```{r}
library(ggmap)
library(ggplot2)
# Mapping top 10 poor performance & 1-2 star states
states <- map_data("state")
poorstates <- c("florida","texas","georgia","california","new york","ohio","illinois","tennessee","louisiana","michigan","pennsylvania","maryland")
poorstates1 <- states %>% filter (region %in% poorstates)
ggmap(usbasemap) + geom_polygon(data = poorstates1, aes(x=long, y=lat, group = group), fill = NA, color = "red") + ggtitle("Top 10 poor performance & 1-2 star rated states")

ggmap(usbasemap) + geom_polygon(data = poorstates1, aes(x=long, y=lat, group = group), fill = "yellow", color = "red")
```
# Mapping prevalence
```{r}
library(ggmap)
library(ggplot2)
prev <- ggmap(usbasemap) + geom_point(aes(x=longitude, y=latitude, color=`avg(DataValue)`), data=alldata, alpha=.5, na.rm = T)  + scale_color_gradient(low="yellow", high="blue") + ggtitle("Chronic Kidney Disease Prevalence in US States")
plot(prev)
```
# Kmeans clustering
```{r}
# Create data frame "allfac.num4" with variables for Kmeans cluster analysis
# Filtering by mortality value outliers based on the distribution of Mortality
allfac.num4 <- alldata %>% filter(mortality_rate_facility >= 5 & mortality_rate_facility <= 30) %>% select(provider_number,network,chain_owned,five_star,perfcats, mortality_rate_facility,readmission_rate_facility, strr_measure_score,standardized_hospitalization_ratio, standardized_readmission_ratio,total_performance_score, nhsn_bsi_measure_score,anemia_management_measure_score,  kt_v_dialysis_adequacy_combined_measure_score,  vascular_access_combined_measure_score, mineral_metabolism_measure_score,X_of_dialysis_stations) %>% filter(complete.cases(.)) 

allfacnum4 <- allfac.num4 %>% select_if(is.numeric)
scaled.allfac <- scale(allfacnum4)

# Create a list 'kvalue' to store results of each kmeans cluster algorithm model
kvalue <- list()
for(i in 1:8){
  kvalue[[i]] <- kmeans(scaled.allfac,i)
}

# Calculating ratio between sum of squares and total sum of squares
bss_totss <- list()
for (i in 1:8){
  bss_totss[[i]] <- kvalue[[i]]$betweenss/kvalue[[i]]$totss
}
# Check the "elbow" plot
plot(1:8,type="b",bss_totss)

# Scatterplots & centroids for mortality vs readmission rates by cluster
for (i in 1:8){
  plot(allfacnum4$mortality_rate_facility,allfacnum4$total_performance_score,
       col=kvalue[[i]]$cluster)
       points(kvalue[[i]]$centers[,c(1,2)],col=1:8,pch=23,cex=4)
}

# Examining 5 clusters
clustnum <- factor(kvalue[[5]]$cluster)
allfac.num4$clust <- clustnum
allfac.num4 %>% ggplot(aes(x=total_performance_score)) + geom_histogram() + 
ggtitle("Total performance score by clusters") + facet_wrap(~clust)

# Examining 4 clusters
clustnum <- factor(kvalue[[4]]$cluster)
allfac.num4$clust <- clustnum
allfac.num4 %>% ggplot(aes(x=mortality_rate_facility)) + geom_histogram() + 
  ggtitle("Mortality rates by clusters") + facet_wrap(~clust)


ggplot(allfac.num4,aes(allfac.num4,x=allfac.num4$mortality_rate_facility,
                     y=allfac.num4$readmission_rate_facility,
                     color=allfac.num4$clust)) + geom_point() + facet_wrap(~allfac.num4$perfcats)

# Using best cluster estimate for 3 clusters
clustnum <- factor(kvalue[[3]]$cluster)
allfac.num4$clust <- clustnum
allfac.num4 %>% ggplot(aes(x=standardized_hospitalization_ratio)) + geom_histogram() + 
ggtitle("Standardized hospitalization ratio by 3 clusters") + facet_wrap(~clust)

plot(allfacnum4$mortality_rate_facility,allfac.num4$readmission_rate_facility,
     col=kvalue[[3]]$cluster,pch=19)

# Using best cluster estimate for 7 clusters
clustnum <- factor(kvalue[[7]]$cluster)
allfac.num4$clust <- clustnum
allfac.num4 %>% ggplot(aes(x=standardized_hospitalization_ratio)) + geom_histogram() + 
ggtitle("Standardized hospitalization ratio by 7 clusters") + facet_wrap(~clust)

```
# Heirarchical clustering
```{r}
library(NbClust)
library(mclust)
# Use scaled dataset from above
dist.allfac <- dist(scaled.allfac)
allfac.clust <- hclust(dist.allfac,method="complete")
plot(allfac.clust,cex=0.1)
rect.hclust(allfac.clust, k=5, border="red")
```

```{r}
# Checking the groupings for 2-5 clusters
g45 <- cutree(allfac.clust, k = c(4,5))
table(grp4 = g45[,"4"], grp5 = g45[,"5"])
# clusters 2 thru 5
g23 <- cutree(allfac.clust,k=c(2,3))
table(grp2 = g23[,"2"], grp3 = g23[,"3"])
```
```{r}
library(ggplot2)
allfac_5 <- cutree(allfac.clust,5)
clust_labs <- factor(allfac_5)
allfac.num4$hlabs <- clust_labs
allfac.num4 %>% ggplot(aes(x=total_performance_score)) + geom_histogram() + 
ggtitle("Total performance score by clusters") + facet_wrap(~hlabs)

# Check the fit
table(allfac.num4$five_star,allfac.num4$hlabs)
```
# Using R packages NbClust to predict number of clusters
```{r}
library(NbClust)
res2 <- NbClust(scaled.allfac,distance="euclidean",min.nc=2,
                max.nc=8,method="complete",index="all")
```
# Package mclust
```{r}
library(mclust)
allfac.mclust <- Mclust(scaled.allfac)
allfac.mclust
#'Mclust' model object:best model: ellipsoidal, equal volume and 
# shape (EEV) with 7 components
```
# Regression analysis
```{r}
library(dplyr)
# Create data frame "allfac.mod1" with variables for regression analysis
allfac.mod1 <- alldata %>% filter(mortality_rate_facility >= 5 & mortality_rate_facility <= 30) %>% select(provider_number,network,chain_owned,five_star,perfcats, mortality_rate_facility,readmission_rate_facility, strr_measure_score,standardized_hospitalization_ratio, standardized_readmission_ratio,total_performance_score, nhsn_bsi_measure_score,anemia_management_measure_score, kt_v_dialysis_adequacy_combined_measure_score,vascular_access_combined_measure_score,mineral_metabolism_measure_score,X_of_dialysis_stations,  rdsallty4_f,`avg(DataValue)`,rac1y4_f,rac2y4_f,rac3y4_f,rac4y4_f,rac5y4_f,eth1y4_f,eth2y4_f,eth3y4_f,dis1y4_f,dis2y4_f,dis3y4_f,dis4y4_f,rdsallty4_f)                                                                              
# Selecting complete numeric cases
allfacnum5 <- allfac.mod1 %>% filter(complete.cases(.)) %>% select_if(is.numeric)
```
# Correlations
```{r}
library(corrplot)
cor.allfacrates <- cor(allfacnum5)
corrplot(cor.allfacrates,tl.cex=0.5)
```
# Calculating correlations
```{r}
library(dplyr)
# Function 'corfun' to calculate correlations
# Inputs: df=dataframe, a=variable1, b=variable2
corfun <- function(df=allfacnum5,a,b){df %>% summarize(N=n(),r=cor(a,b))}

# Checking correlations with mortality rate
corfun(a=allfacnum5$mortality_rate_facility,b=allfacnum5$standardized_readmission_ratio)
corfun(a=allfacnum5$mortality_rate_facility,b=allfacnum5$standardized_hospitalization_ratio)
corfun(a=allfacnum5$strr_measure_score,b=allfacnum5$mortality_rate_facility)
# Correlations with Mortality are very low, checking steps function output 
allfac.MR <- lm(mortality_rate_facility~.,data=allfacnum5)
steps.MR <- step(allfac.MR,direction="backward",data=allfacnum5)
summary(steps.MR)
# add leaps
library(leaps)
leaps <- regsubsets(mortality_rate_facility~.,data=allfacnum5)
leaps
plot(leaps,scale="r2")
```
# Creating scatterplots 
```{r}
library(ggplot2)
allfacnum5 %>% ggplot(aes(x = total_performance_score, y = readmission_rate_facility)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)

allfacnum5 %>% ggplot(aes(x = total_performance_score, y = standardized_hospitalization_ratio)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) 

```
# Modelling total performance scores
```{r}
# Consult plot to idenitfy variables related to total performance score
corfun(a=allfacnum5$total_performance_score,b=allfacnum5$readmission_rate_facility)
# Model 1: Creating a model for total performance score as dependent variabe, readmission rate as a predictor
mod1 <- lm(total_performance_score ~ readmission_rate_facility, data=allfacnum5)
summary(mod1)
```
```{r}
# Model 2: Adding Transfusion score to model 1
corfun(a=allfacnum5$total_performance_score,b=allfacnum5$strr_measure_score)
mod2 <- lm(total_performance_score ~ readmission_rate_facility + strr_measure_score, data=allfacnum5)
summary(mod2)
```
```{r}
# Model 3: Adding Infection measure score
corfun(a=allfacnum5$total_performance_score,b=allfacnum5$nhsn_bsi_measure_score)
mod3 <- lm(total_performance_score ~ readmission_rate_facility + strr_measure_score + nhsn_bsi_measure_score, data=allfacnum5)
summary(mod3)
```
```{r}
# Model 4: Adding dialysis adequacy measure score
mod4 <- lm(total_performance_score ~ readmission_rate_facility + strr_measure_score + nhsn_bsi_measure_score + kt_v_dialysis_adequacy_combined_measure_score, data=allfacnum5)
summary(mod4)
```
```{r}
# Model 5: Adding vascular access score
mod5 <- lm(total_performance_score ~ readmission_rate_facility + strr_measure_score + nhsn_bsi_measure_score + kt_v_dialysis_adequacy_combined_measure_score + vascular_access_combined_measure_score, data=allfacnum5)
summary(mod5)
```
```{r}
#Model 6: Adding standardized hospitalization ratio
mod6 <- lm(total_performance_score ~ readmission_rate_facility + strr_measure_score + nhsn_bsi_measure_score + kt_v_dialysis_adequacy_combined_measure_score + vascular_access_combined_measure_score + standardized_hospitalization_ratio, data=allfacnum5)
summary(mod6) 
```
```{r}
# Regression plots 
hist(mod6$residuals)
qqnorm(mod6$residuals)
qqline(mod6$residuals)
plot(mod6$fitted.values,mod6$residuals)
```
# Modeling total performance score by using lm and steps
```{r}
psmodel <- lm(total_performance_score ~ .,data=allfacnum5)
steps.PS <- step(psmodel,direction="backward",data=allfacnum5)
summary(steps.PS)
```
# Sentiment Analysis: Rvest and tidytext
```{r}
library(rvest)
library(tidytext)
library(dplyr)
# Dialysis support groups text analysis using tidytext
# Discussion forum on Kidney failure and Chronic Kidney Disease
url <- "https://patient.info/forums/discuss/browse/kidney-failure-and-ckd-1300#"
webtext <- read_html(url) %>% 
  html_nodes(xpath='//*[(@id = "group-discussions")]//p') %>% html_text()

# Create data frame with 35 rows based on text file
text_df <- data_frame(line = 1:35, text = webtext)
# use "unnest tokens function"
tidyforum <- text_df %>% unnest_tokens(word,text)
# load stopwords data
data("stop_words")
# remove stopwords
tidyforum <- tidyforum %>% anti_join(stop_words)
tidyforum %>% count(word, sort = TRUE)
# load sentiments data
data("sentiments")
forumsentiment <- tidyforum %>% inner_join(get_sentiments("bing"))
forumsentiment %>% count(sentiment)
webtext[1:3]
```
# Sentiment analysis: Package Rtimes
```{r}
library(rtimes)
library(tm)
library(wordcloud)
library(SentimentAnalysis)

Sys.setenv(NYTIMES_AS_KEY = "1f3c87aaf52e46129e7525843d872aae")
articles <- as_search("dialysis",begin_date = 20150101,end_date = 20180401, fl = c('web_url','headline','snippet'),all_results = TRUE)

snippetlist <- list(articles$data$snippet)
snippetlist_text <-  sapply(snippetlist,function(x) gettext(x))

Corpus(VectorSource(snippetlist_text))
dia_articles <- Corpus(VectorSource(snippetlist_text))
inspect(dia_articles[1])
dia_articles <- tm_map(dia_articles,stripWhitespace)
dia_articles <- tm_map(dia_articles,removeNumbers)
dia_articles <- tm_map(dia_articles,content_transformer(tolower))
dia_articles <- tm_map(dia_articles,removeWords,stopwords("en"))
dia_articles_clean <- tm_map(dia_articles,removePunctuation)

library(wordcloud)
wordcloud(dia_articles_clean)
wordcloud(dia_articles_clean,scale = c(2,0.5),colors = rainbow(50))
dia_articles_tdm <- TermDocumentMatrix(dia_articles_clean)
dia_articles_tdm1 <- dia_articles_tdm
dia_articles_tdm_m <- as.matrix(dia_articles_tdm)
dia_articles_freq <- rowSums(dia_articles_tdm_m)
diaterms_freq_sorted <- sort(dia_articles_freq,decreasing=T)
article_sent <- analyzeSentiment(dia_articles_tdm1)
str(article_sent)
binaryvalues <- convertToBinaryResponse(article_sent$SentimentQDAP)
(binary <- table(binaryvalues))
grep("positive",binaryvalues)[1:20]
snippetlist_text[c(16,20,30)]
```
# Sentiment Analysis: Package Rtweet
```{r}
# Load libraries
library(rtweet)
library(tm)
library(SentimentAnalysis)
library(wordcloud)

# searching for tweets in "poor" performing states
dia <- search_tweets("Dialysis",lang="en",n = 5000,retryonratelimit = TRUE,geocode = lookup_coords("fl","tx","ca","ga","ny"))
dia1 <- plain_tweets(dia$text)
dia1 [1:20]
dia_rawtweets <- dia1
diacleaned_tweets_vec <- VectorSource(dia_rawtweets)
diacleaned_tweets_corp <- VCorpus(diacleaned_tweets_vec)

dia_stopwords <- c(stopwords("english"),"dialysis","patent","kidney","gtgtgt","job","careerarc")
# function tm cleaner from notes
tm_cleaner <- function(corpus, stop=dia_stopwords, rm_num=TRUE) {
  require(tm)
  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, removeNumbers)
  corpus <- tm_map(corpus, content_transformer(tolower))
  corpus <- tm_map(corpus, removeWords, stop)
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, 
                   content_transformer(function(x) gsub("http\\w+", "", x)))
  return(corpus)
}
diacleaned_tweets_corp <- tm_cleaner(diacleaned_tweets_corp,stop=dia_stopwords)
diacleaned_tweets_tdm <- TermDocumentMatrix(diacleaned_tweets_corp)
dialysis_cleaned_tweets_tdm <- diacleaned_tweets_tdm
diacleaned_tweets_tdm_m <- as.matrix(diacleaned_tweets_tdm)
diaterms_freq <- rowSums(diacleaned_tweets_tdm_m)
diaterms_freq_sorted <- sort(diaterms_freq,decreasing=T)
diaterms_freq_sorted[1:10]
diadf <- data.frame(word=names(diaterms_freq_sorted),
                    freq=diaterms_freq_sorted,stringsAsFactors = FALSE)

# Create wordcloud
wordcloud(diadf$word,diadf$freq,scale = c(2,0.5),colors = rainbow(50))
# perform sentiment analysis
diatweetsentiment <- analyzeSentiment(diacleaned_tweets_tdm)
binaryvalues <- convertToBinaryResponse(diatweetsentiment$SentimentQDAP)
(binary <- table(binaryvalues))
grep("positive",binaryvalues)[1:15]
dia1[c(3,10,19)]
```
# Patient Utility Function
```{r}

c <- c("10461")
facinfo <- function (c) {
#
# FUNCTION TO PROVIDE FACILITY ADDRESS, RATING INFORMATION & MAP 
#  
# INPUTS: 
#
# C - Character vector with one or more zipcode values
# 
# OUTPUTS: A dataframe with facility address, network, and patient rating
# Load libraries
library(dplyr)
library(leaflet)
# Create empty list 'infolist'
infolist <-list()
# Create dataframe 'tmpzip' with filtered data corresponding to 'c'zip code values 
tmpzip <- alldata %>% filter(zip %in% c)
# For loop to perform functions on each row value of tmpzip
for (row in 1:nrow(tmpzip)){
# Create & populate 'infolist' elements from corresponding 'tmpzip' elements
  
infolist$Name <- tmpzip$facility_name
infolist$Address <- tmpzip$address_line_1
infolist$City <- tmpzip$city
infolist$State <- tmpzip$state
#infolist$Contact <- tmpzip$Pho
infolist$Network <- tmpzip$network
infolist$Rating <- tmpzip$facrating
infolist$Score <- tmpzip$perfcats}
# Create 'listout' with Adress, phone, Network, Rating, Score information for each facility
listout <- list(Address=paste(infolist$Name,infolist$Address,infolist$City,infolist$State,sep=","),
                Network=infolist$Network,
                "Patient Rating"=infolist$Rating,
                "Performance Score"=infolist$Score)
# Create output 'retdf' dataframe from 'listout' elements 
retdf <- data.frame (listout)
# Create Leaflet map 'locations' by using lat-long values for each facility
locations <- leaflet(tmpzip) %>% addTiles() %>% 
    addMarkers(~longitude,~latitude, popup=~facility_name,
               clusterOptions=markerClusterOptions())
# Create & populate final output list 'facinfoloc' elements with 'retdf' and 'locations'
facinfoloc <- list("Facility Information"=retdf,Map=locations)
# Provide dataframe and map outputs
return(facinfoloc)
}
```
```{r}
facinfo(c)
```

```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
